---
title: "Enrichment Analysis Visualization"
shorttitle: "KYCG-VISUAL"
package: knowYourCG
output: rmarkdown::html_vignette
fig_width: 6
fig_height: 5
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{"4. Enrichment Visualization"}
  %\VignetteEncoding{UTF-8}
---

# Visualize Enrichment Results

There are many ways one can visualize enrichment results. Here we
demonstrate a few popular ones supported in sesame. Dot plot:

```{r ky14, echo=TRUE, message=FALSE}
library(SummarizedExperiment)
library(sesame)
library(knowYourCG)
sesameDataCache("MM285.tissueSignature")
kycgDataCache()
df <- rowData(sesameDataGet('MM285.tissueSignature'))
query <- df$Probe_ID[df$branch == "fetal_liver" & df$type == "Hypo"]
results <- testEnrichment(query, "TFBS")
```

```{r ky15, fig.width=5, fig.height=3.5}
KYCG_plotDot(results)
```

or a bar plot:

```{r ky16, fig.width=6.5, fig.height=4, message=FALSE}
library(ggplot2)
library(wheatmap)
p1 <- KYCG_plotBar(results, label=TRUE)
p2 <- KYCG_plotBar(results, y="estimate") + ylab("log2(Odds Ratio)") +
    xlab("") + theme(axis.text.y = element_blank())
WGG(p1) + WGG(p2, RightOf(width=0.5, pad=0))
```
 
or a volcano plot. Here we need a two-tailed test to show both enrichment and
depletion (the default is one-tailed):

```{r ky17, fig.width=6, fig.height=4, warning=FALSE, message=FALSE}
query <- df$Probe_ID[df$branch == "fetal_brain" & df$type == "Hypo"]
results_2tailed <- testEnrichment(query, "TFBS", alternative = "two.sided")
KYCG_plotVolcano(results_2tailed)
```

and a Waterfall plot:

```{r ky18, fig.width=6, fig.height=4.5}
KYCG_plotWaterfall(results)
```

If you have a list of cg groups and they were tested against the same set of
databases, you can make a point range plot to summarize the overall trend:

```{r ky19, fig.width=3.5, fig.height=4}
## pick some big TFBS-overlapping CpG groups
cg_lists <- getDBs("MM285.TFBS", silent=TRUE)
queries <- cg_lists[(sapply(cg_lists, length) > 40000)]
res <- lapply(queries, function(q) {
    testEnrichment(q, "MM285.chromHMM", silent=TRUE)})

## note the input of the function is a list of testEnrichment outputs.
KYCG_plotPointRange(res)
```

plot enrichment over metagene:

KnowYourCG builds in several metagene/meta-feature coordinates. One can test
enrichment over meta genes or simply plot the mean over metagenes:

```{r ky20, plot-meta1, fig.width=5, fig.height=4, message=FALSE}
sdf <- sesameDataGet("EPIC.1.SigDF")
KYCG_plotMeta(getBetas(sdf))
```

Here we picked some transcription factor binding sites-overlapping CpGs and
tested against chromHMM states. As expected, most of these CGs are enriched at
promoters and enhancers but depleted at heterchromatic regions.

## Manhattan plot

Manhattan plots are an intuitive way to visualize the results from large scale
"omics" studies that investigate genetic or epigenetic features on genome wide
scales across large groups of samples. Here we demonstrate how the
`KYCG_plotManhattan()` function can be used to visualize the chromosomal
distribution and significance level of CpG probes from an EWAS
study. `KYCG_plotManhattan()` takes a named numeric vector with CpG probeIDs as
names and numeric values (P,Q,logFC,etc) obtained from analysis as values. To
plot EWAS results, simply load the necessary libraries and pass the named
numeric vector to `KYCG_plotManhattan()`

By default, `KYCG_plotManhattan()` will infer the array platform from the
probeIDs and retrieve the corresponding GRanges object to obtain probe
coordinates. However, the platform and GRanges objects can be pre - specified
if offline. Color, title, and the threshold to label significant probes on the
plot can also be specified:

```{r ky21, fig.width=6, fig.height=4, message=FALSE}
library(ggrepel)
smry_pvals = readRDS(url("https://github.com/zhou-lab/InfiniumAnnotationV1/raw/main/Test/20220413_testManhattan.rds"))
KYCG_plotManhattan(-log10(smry_pvals), platform="HM450",
    title="Differentially Methylated CpGs - EWAS Aging",
    col=c("navy","skyblue"), ylabel = bquote(-log[10](P~value)), label_min=30) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Feature Aggregation

In addition to hypothesis testing, knowYourCG also uses the curated database
sets for feature engineering. We have a pre-curated summarized experiment
containing a samplesheet and beta value matrix corresponding to about 467 MM285
samples with 20k probes. The samplesheet includes UIDs pertaining to the sample
and several categorical/numerical features. To use this data for a linear
model, we will extract the most relevant prevalent features.

``` {r ky22, echo=TRUE, message=FALSE}
library(SummarizedExperiment)
betas = assay(sesameDataGet('MM285.467.SE.tissue20Kprobes'))
```

We have found that it is computationally expensive to perform a linear
model/generalized linear model on a feature set of individual
CpGs. Additionally, interpreting the mechanism the significantly contributing
CpGs is non-trivial due to their complex interactions. We hope to leverage
these pre-curated database sets by using their beta value summary statistics as
features instead. We will calculate the summary statistics for the betas matrix
using a list of database sets. The default is to calculate the mean.

```{r ky25}
stats <- dbStats(betas, 'MM285.chromHMM')
head(stats[, 1:5])
```

Just from the few database set means above, we can see that TSS are
consistently hypomethylated, which is consistent with known biology.

```{r ky26, fig.width=6, fig.height=6}
library(wheatmap)
WHeatmap(both.cluster(stats)$mat, xticklabels=TRUE,
    cmp=CMPar(stop.points=c("blue","yellow")))
```

## Feature Annotation

We can also use KYCG database to annotate an arbitrary probe set. This can be
done using the `KYCG_annoProbes` function:

```{r ky30}
query <- names(sesameData_getManifestGRanges("MM285"))
anno <- annoProbes(query, "designGroup", silent = TRUE)
head(anno)
```

# Session Info

```{r session-info}
sessionInfo()
```
